### 2018-05-17 revision/reply_to_revs_2.tex
First draft of replies

\documentclass[]{article}

\usepackage{graphicx}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

%\maketitle

\section*{Response to Reviewer 1}

We thank the reviewer for his/her constructive comments and are pleased with the positive assessment or our work. We have addressed the specific comments made by the reviewer as described below. We hope that the revised manuscript can now be accepted for publication.

\dotfill

\textbf{1. Reviewer:} \textit{One comparison that came to mind reading the paper, is that the results show very little value in optimizing the most-downstream turbines, and in general, improvement in power comes from modifications upstream, and that self-optimization is not possible. This stands in contrast to a result such as:}

\textit{Ciri, Umberto, Mario Rotea, Christian Santoni, and Stefano Leonardi. “Large Eddy Simulation for an Array of Turbines with Extremum Seeking Control.” In American Control Conference. Boston, MA, 2016. }

\textit{Where the TSR of downstream turbines is re-optimized for wake conditions (and the upstream turbine is left as is at the end of the optimization). It would seem the difference in modeling methods and/or how turbine control is implemented yields the different results, but I believe it would be worth discussing the difference, for example around the paragraph beginning with "The figure shows that the first row (R1)..." on page 12.}

\textbf{Response:} The reason for this different behavior of the downstream vs upstream turbines is twofold:
\begin{enumerate}
	\item The Extremum Seeking Control (ESC) approach by Ciri \emph{et al.} differs fundamentally from our approach. In our approach, we simultaneously optimize all control variables over a given time horizon, which allows for full dynamic cooperation and interaction between all turbines. In the ESC, controls are optimized one by one in the sense that ESC is initiated on the first row turbine and will be activated on the subsequent row only after its upstream neighbor has converged to a steady state solution. In this way, dynamic control and interaction between controls in different turbines are discarded a priori. Furthermore, although this is not an inherent limitation of the ESC approach, Ciri \emph{et al.} optimize for individual turbine power instead of total wind-farm power (see Eq. 10 $P_i$ is used instead of $\sum_i P_i$). This is explicitly mentioned in the title of their follow up paper (Ciri \emph{et al.} Large-eddy simulations with extremum-seeking control for individual wind turbine power optimization)

	\item In our modeling approach, we make abstraction of detailed turbine dynamics in the sense that we directly control the thrust coefficient $C_T'$ instead of the generator torque or tip speed ratio. In this sense, the optimal thrust coefficient value of a turbine with no downstream neighbors can be analytically found to be $C_T' = 2$. In Ciri \emph{et al.}, the authors indicate that the reason why the TSR in the downstream rows (around 6.5) does not correspond to the individual optimal TSR (7.5) of individual turbines is that the TSR depends on a reference velocity and (direct quotation) ``As a consequence, the TSR is not a meaningful indicated of a waked turbine operating condition".

\end{enumerate}

Summarizing, the inverse behavior of the control dynamics compared to Ciri \emph{et al.} can be attributed to the difference in approach to wind-farm control (dynamic full-farm optimization vs steady sequential and individual optimization) as well as the implementation of the turbine control ($C_T'$ vs $\lambda$). In our opinion, the current study (full-farm dynamic wind-farm optimization) and that of Ciri \emph{et al.} (providing a practical ESC method for optimizing individual turbines) cannot be directly compared.

However, we understand the contrast in the results and therefore added a small comment to the revised manuscript  (l. ??, p. ??):

``
... is not monotonous. \textbf{Note that these observations are in contrast to the study of Ciri et al. 2017, where extremum seeking control is applied to optimize individual turbines in an array of three turbines. They found that set points in the first-row turbines remain untouched from their locally optimal values whereas downstream turbines are re-optimized for wake conditions. This contrast can be attributed to our dynamic and simultaneous wind-farm control approach, which differs from their sequential individual set-point control.  Furthermore, self-optimization in the latter study is achieved by recalibrating the torque gain. The baseline controller in our current actuator disk model makes abstraction of pitch and generator torque control and keeps $C_T'$ at its theoretical optimum instead.}
''

WM - WEET NIET OF WE DIT WEL ECHT MOETEN DOEN, MISSCHIEN GEWOON REVIEWER OVERTUIGEN DAT WE DIT NIET INCLUDEN? TE BESPREKEN

\dotfill

\textbf{2. Reviewer:} \textit{Figure 3/related text: Would another way to describe Ct2 vs Ct3 be that Ct2 can only lower the thrust, while Ct3 is allowed to raise it?}

\textbf{Response:} Indeed, this is correct. We have mentioned this explicitly in the revised manuscript as follows (p. ??, l. ??):

``
... and the maximal thrust coefficient $C_{T,\rm max}' = 2$ or 3, \textbf{with thrust forces that can respectively only be reduced (underinductive), or also increased (overinductive) compared to the Betz optimum at $C_T' = 2$} (see Eq. 5). 
''

\dotfill

\textbf{3. Reviewer:} \textit{``... NREL 5MW rotor with a 50\% increase in chord length ...'' does this imply the method is currently assuming the chord length is variable? Could this not be achieved by a change in pitch angle?}

\textbf{Response:} No, the method does not imply a variable chord length.

Current turbines are designed to approach maximum $C_T'$ values around 2, corresponding to the Betz limit. Although this maximum value can be increased by increasing the operational tip speed ratio (TSR), aiming to do so by adapting the pitch angle would inevitably lead to severe efficiency losses due to stall on the turbine blades. Therefore, we provide an example of how an alternative turbine design (i.e. with an increased chord length and operational TSR) could attain a maximum $C_T'$ of 3.5. Given such a turbine design, achieving thrust ratings $C_T' < 3.5$ is straightforward by pitching blades towards the feather position.

We have slightly modified the statement in the revised manuscript to avoid any confusion with regard to possibly having the chord length as a control variable. (p. ??, l. ??):

``
... be \textbf{attained by the NREL 5MW turbine with a slight modification in the rotor design}, i.e. a 50\% increase in blade chord length and an operational tip speed ratio 25\% higher than 
the original design value (see Goit and Meyers, 2015, Appendix A). \textbf{Furthermore, given this redesign, dynamic reductions from this maximum value could be realized through pitch control. More specifically, based on speculated specifications of similar turbines in practice, the NREL 5 MW is designed with a pitch rate of $8^{\circ}/$s. This implies that the rotor can go from power maximization ($0^\circ$ pitch) to total curtailment in fully feathered blade position ($90^\circ$ pitch) in 11.25 seconds (Jonkman et al. 2009).}
''

\clearpage
\section*{Response to Reviewer 2}
We thank reviewer 2 for reading our work and for providing detailed feedback which has improved the quality of the manuscript. We are pleased with his/her positive assessment of our research and have addressed the specific comments in the manuscript as described below. We hope that the revised manuscript can now be accepted for publication.

\dotfill

\textbf{Reviewer: } \textit{...LES setup is described/illustrated properly - except of the characteristics of the turbines/actuator disks considered. Would be nice to have the size of the disks explicitly noted (as they are somewhat hidden in Figure 2) to have a more clear scale of the considered wind farm.}

\textbf{Response: } Indeed, we seem to have overlooked specifying the exact turbine dimensions in Section 2.2. We have updated the manuscript as follows (l XX p YY):

``
... 12 rows by 6 columns. \textbf{The wind turbines have a hub height $z_h = 100$~m with a rotor diameter $D = 100$~m, and are spaced apart by 6$D$ in both axial and transversal directions.}
''

\dotfill

\textbf{Reviewer: } \textit{While describing the case setup in Section 2.2, the "flow advancement time", $T_A$ (also
	referred in Figure 1) is considered as half of the prediction horizon $T$. Would $T_A$ (and
	therefore $T$) be inflow dependent as the time delay (the time it takes for particles to
	move from the upstream to downstream turbine(s))? Have you investigated if changing
	$T$ (and/or $T_A$) has any effects on the resulting optimum $C_T$ set-points and on the power
	gain?}

\textbf{Response: } Indeed, $T$ and $T_A$ are important parameters in the receding-horizon approach. The optimization horizon $T$ can indeed be considered inflow-dependent. In our case, it is chosen as the time it takes for the flow to pass approximately four rows of turbines. In theory, it is desirable to have $T$ cover an entire wind-farm throughflow. In practice however, control over long time horizons is complicated by the chaotic nature of turbulence accompanied by adjoint gradient inaccuracies, and we choose $T$ as long as possible for our optimization to be unharmed by these inaccuracies (i.e. 240 s). The sensitivity of the power gains to the time horizon has not been formally quantified, yet it is reasonable to expect that the potential for beneficial interaction between turbines that are located more than three rows apart is limited.

The choice of $T_A$ is driven by conflicting incentives of reducing computational cost ($T_A \rightarrow T$) and mitigating finite-horizon effects ($T_A \rightarrow 0$). As a rule of thumb, we have been using $T_A = T/2$ in previous work. In the MM17 paper we have also tried $T_A = T/4$ and found that the effect on power gain was limited. (TODO: Perhaps include a figure here? )

Both concepts described above are discussed in more detail in the MM17 paper. To keep the description of the case setup concise in the current work, we simply add that $T = 240$ s corresponds to the time to pass four turbine rows, and further refer to MM17 for further elaboration of methodological choices. We have updated the revised manuscript as follows (l XX p XX):

``
... with a prediction horizon $T = 240$ s \textbf{(i.e. the time it takes for the flow to pass four rows of turbines)} and a flow ...
'',

and (l XX p XX, end of Section 2.2):

`` 
... and $\tau = 30$ s. \textbf{The choice of (and sensitivity to) setup parameters is further elaborated in MM17.}
''

\dotfill

\textbf{Reviewer: } \textit{As clearly seen in Figure 4c and 4d, there is a significant increase in turbulence
	further downstream. In addition to the TKE and the transport, would be nice to have the
	turbulence intensity TI values (as listed later on page 20, 10\% for the baseline case),
	both for the baseline case and the maximum added TI reported - possibly somewhere
	around Figure 4. That again would give an indication on the applicability compared to
	the field values observed. Also note the typo in the caption of Figure 4: after c) all the
	subplots are marked to be continuously c).}

\textbf{Response: }

We have added the requested TI values in the discussion around Figure 4 as follows (p ?? l ??):

``
.., for which an enhanced recovery was found as discussed above. \textbf{The turbulence intensity $TI \equiv (2k/3  )^{1/2}/(U_x^2+U_y^2+U_z^2)^{1/2}$ at hub height (not shown in the figure) is 10\% at the inlet for both the reference case and the controlled case. The combination of reduced near-wake mean velocities and increased velocity fluctuations in the controlled case increase local $TI$ in the turbine wakes (ranging from $\approx 2$\%-points in the wakes of middle rows to $\approx 12$\%-points in the first and last rows). This increase in turbulence intensity dissipates to below $1$\%-point difference at 10$D$ downstream of the last row.}
''

Furthermore, we thank the reviewer for pointing out the typos in the caption. This has been fixed in the revised manuscript.

\dotfill

\textbf{Reviewer: } \textit{On page 10, around line 10, the argument of "upstream actions do not require a
	specific downstream response in order to increase power in that downstream row",
	which is also paraphrased in the conclusions, needs to be elaborated. This rather
	broad conclusion seem to oversee the probability of the curtailment of the downstream
	turbine where down-regulation might be inevitable for certain CT set-points assigned
	to downstream turbine(s) in the resulting optimization. Could be partially true for the
	investigated C3t5 case since there observed very limited curtailment even at the most
	upstream turbine (as in Figure 3b). However, also seen in Figure 8b (except of the
	very last row as the authors indicated), there seem to be still a difference between on
	the power gain at turbine R11 for the scenarios of R1-R10 and R1-R11. Narrowing the argument to the considered case or very little to no downstream curtailment CT
	distributions is suggested.}

\textbf{Response: } WM - Ik vind de commentaar van de reviewer hier vrij verwarrend. Weet niet exact wat hij hier bedoelt. Hopelijk is onderstaand antwoord voldoende.

%We thank the reviewer for this comment. We believe that the statement referred to by the reviewer still holds, i.e.
%
%``An interesting observation from this figure is that , for any row $i$ except the last one, the power potential as observed in case C3t5 is almost fully recovered by
%only applying the optimal controls up to row $i-1$''. 
%
%This is a simple observation that stems directly from the data in the figure. Unless we misinterpret the point of the reviewer, we do not agree that a curtailment strategy would invalidate this statement. If, let's say, a curtailment strategy would have been incorporated into the control law for row $i$ (i.e. to the benefit of row $i+1$), the application of the control R1-R$_{i-1}$ would yield even greater power gains in row $i$ than when applying controls R1-R$_i$, so the statement would still hold. 
%
%However, we do agree that statements and conclusions made throughout section 3 are specific to the current case and should therefore be interpreted with care. We have updated the manuscript to be more explicit that results are not overall conclusions for wind-farm control in general, but should be interpreted as observations of the current optimal control case. This has been done at the beginning of Section 3, prior to the elaboration of analyses and experiments as follows (l XX p XX):
%
%`` 
%... to uncover some of the characteristics of these control signals. \textbf{Note that the conclusions drawn within this paragraph should be interpreted as observations of the current C3t5 optimal control cases, given specific wind-farm layout and flow conditions, and hence cannot just be generalized for any wind-farm control in general.}
%''
%
%We hope that the updates to the manuscript encompass the concerns expressed by the reviewer.

We thank the reviewer for this comment. We agree that statements and conclusions made throughout section 3 are specific to the current case and should therefore be interpreted with care. However, possible curtailment is not the only reason why certain statements in this section could be invalidated, e.g. possible phase-locked correlated control actions between turbines (which were found not to be present in the current case) would invalidate the same statement as referred to by the reviewer.

We have updated the manuscript to be more explicit in the sense that results are not overall conclusions for wind-farm control in general, but should be interpreted as observations of the current optimal control case. This has been done at the beginning of Section 3, prior to the elaboration of analyses and experiments as follows (l XX p XX):

`` 
... to uncover some of the characteristics of these control signals. \textbf{Note that the conclusions drawn within this section should be interpreted as observations of the current C3t5 optimal control cases, given specific wind-farm layout and flow conditions, and hence cannot just be generalized for any wind-farm control in general.}
''

However, we do believe that the sinusoidal thrust control strategy is something that can be generalized to other wind-farm setups (as evidenced by Section 4.2.2). We hope that the updates to the manuscript encompass the concerns expressed by the reviewer.

\dotfill

\textbf{Reviewer: } \textit{On page 13, line 13, "the presence of the flow invariant features of the control signals"
	needs further justification as Figure 11 would also depend on how variant the flow
	features are in the simulations. That should include both the spatial and temporal
	variance within the 30-min window. As far as the field measurements are concerned,
	high spatial and temporal correlations are observed. For the former, Figure 2 gives a
	brief idea about the wind speed range between the columns, that can be referred here.
	For the latter, time series or relevant temporal statistics can be presented to assess the
	randomness and strengthen the hypothesis.}

\textbf{Response: } We thank the reviewer for this useful comment. We have quantified spatial variance (between columns) as well as temporal variance (between time windows) of the flow field by investigating the correlation between the incoming disk-averaged velocity at 6$D$ upstream of the first-row turbines (illustrated for two columns of the wind farm in Figure \ref{fig:upstream_vel} of the current document). As shown in the Figure, there is considerable variation of the incoming velocities between columns and between time windows, and velocity fluctuations seem qualitatively uncorrelated. To keep the discussion in the manuscript concise, we do not include the figure below but instead report the average Pearson correlation coefficient between the incoming velocity fluctuations in different columns (for the full time horizon) and in different time windows (averaged over all columns), with low values of 0.12 and 0.07 respectively.

We have incorporated this in the revised manuscript as follows (p ??):

``
... the column swap is performed in 2 random independent ways. \textbf{The variability of flow conditions for different columns can be qualitatively observed in Figure 2. To further strengthen the hypothesis of the current experiment, we verified that the  correlation between flow conditions in different columns is small, i.e. with an average Pearson correlation coefficient of 0.12 between columns for the incoming velocity fluctuations 6$D$ upstream of the first row.}
''

and (p ??):

``
whereas the time synchronization of control actions to specific flow events is 10 eliminated. Similar to the first case, this is done in 2 random ways, \textbf{and the limited correlation between velocity fluctuations in different time windows was quantified at 0.07}. Figure 11b illustrates the row-averaged power for these...
cases.
''

\begin{figure}
	\includegraphics[width=\textwidth]{upstream_vel.eps}
	\caption{Disk-averaged axial velocity upstream of the first-row turbines for column 1 and column 2 in Figure 2 of the manuscript. Vertical dotted lines demarcate the optimization window boundaries (i.e. at integer multiples of $T_A$) \label{fig:upstream_vel}}
\end{figure}

\dotfill

\textbf{Reviewer: } \textit{On page 17, around line 5, a very nice example on how to implement the optimized
	sinusoidal CT is presented. The practical examples can be further improved by a
	short discussion on the expected response time of such increases in tip speed ratio
	on a machine with high inertia. That would put the estimated sine wave period into
	perspective as well.}

\textbf{Response: } Indeed, the high inertia of the rotor could be a compromise to achieving the desired sinusoidal thrust coefficient behavior. However, given the proposed redesign of the NREL 5MW rotor, dynamic reductions from the maximum $C_T' = 3.5$ could be achieved rapidly by using the fast pitch actuators with which modern turbines are equipped. We have added this comment to the discussion as follows:

``
... be \textbf{attained by the NREL 5MW turbine with a slight modification in the rotor design}, i.e. a 50\% increase in blade chord length and an operational tip speed ratio 25\% higher than 
the original design value (see Goit and Meyers, 2015, Appendix A). \textbf{Furthermore, given this redesign, dynamic reductions from this maximum value could be realized through pitch control. More specifically, based on speculated specifications of similar turbines in practice, the NREL 5 MW is designed with a pitch rate of $8^{\circ}/$s. This implies that the rotor can go from power maximization ($0^\circ$ pitch) to total curtailment in fully feathered blade position ($90^\circ$ pitch) in 11.25 seconds (Jonkman et al. 2009).}
''

\dotfill

\textbf{Reviewer: } \textit{For Section 4.2.3, the header ``Full-scale wind farm test" is a bit misleading... Suggest
	to change to "Full-scale wind farm simulations (in LES)" instead.}

\textbf{Response: } Agreed, we have changed the section header to ``Full-scale wind-farm LES"

\dotfill

\textbf{Reviewer: } \textit{On Figure 19, why would the power decrease after Row 5 for the sinusoidal case?}

\textbf{Response: } This can be explained based on Figure 20: it can be seen from the cross-section views that by sinusoidally varying first-row thrust, the time-averaged axial velocity at the turbine disk for R2 is increased, but the flow above the turbine disk is actually slowed down, hence indicating that momentum is entrained from the internal boundary layer above the turbine canopy. In downstream rows, this high momentum zone expands and mixes with the background flow field. Starting from R5, this zone is almost completely dissipated, and the disk velocity is actually slighter lower than in the reference case. In short, the sinusoidal control actions of the first row actually entrain momentum from the surrounding flow that would otherwise be entrained by natural turbulent mixing further downstream in the wind farm.

Moreover, it can be seen that the full optimal control (C3t5 in Figure 20) however is able to contain the high momentum zone by continuously improving the mixing process with the background boundary layer. The mechanisms thereof however remain elusive to date, and are subject of further investigation.

We have incorporated this discussion in the revised version of the manuscript as follows (p ?? l ??):

``
... with higher rotor velocities in the downstream as well.  \textbf{Note also that, for the fifth row, the disk velocity is slightly lower for the sinusoidal control case than for the reference case, consistent with the decreased power extraction observed in Fig. 19. This can be explained by the fact that first-row control actions cause enhanced entrainment of momentum from the internal boundary layer above the turbine canopy that would otherwise be entrained by natural turbulent mixing in passive downstream rows. In consequence, lesser entrainment occurs for downstream rows, resulting in a slight decrease in disk velocities from the fifth row onwards.}
''

\dotfill

\textbf{Reviewer: } \textit{Page 22 around line 5, the (inevitable) discussions on loads are included. In addition
	to the loads on the controlled upstream turbine, Figure 20(b) indicates partial wakes
	on the further downstream rows of turbines. Therefore, the section should be improved
	by highlighting the probable increase in fatigue loading for not just turbine(s) R1 but for
	the downstream rows as well, possibly starting as early as R3.}

\textbf{Response: } We thank the reviewer for this comment. Agreed, the downstream turbines could also be subjected to increased fatigue loading, and we have included this in the revised manuscript as follows (p ??, l ??):

``
... to fatigue loading \textbf{of the first-row turbines. Furthermore, partial wake alleviation and unsteady passing of abovementioned vortex rings could also increase fatigue loading in downstream rows.} Hence, structural aspects ...
''

\dotfill

\textbf{Reviewer: } \textit{On the grammatical note, the manuscript is clear and easy to follow. The only comment
	might be on the use of Sect. or Section; Fig. or Figure references.}

\textbf{Response: } We have followed the official guidelines for referring to figures and sections as provided by the publisher, paraphrasing from the website:\

\emph{The abbreviation ``Sect.'' should be used when it appears in running text and should be followed by a number unless it comes at the beginning of a sentence.}\\({\small \verb|https://www.wind-energy-science.net/for_authors/manuscript_preparation.html|})

\end{document}